---
title: "Acessing Brazil Central Bank data via API"
format: 
  gfm: default
  pdf:
    pdf-engine: pdflatex
    header-includes: 
    - \usepackage{fvextra}
    - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}, fontsize=\small}
  
    - \usepackage{tcolorbox}
    - \newtcolorbox{functioncard}{
        colback=red!5,
        colframe=red!70!black,
        arc=2mm,
        boxrule=0.8pt}
  html:
    css: stylesheet.css
  
date: "2024-08-05"
editor: source
---


<!------------>
<!-- PART 1 -->
<!------------>
## Brief guidethrough

<p> 
|       This simple library allows the users to acess Brazil Central Bank (BACEN) data by acessing its API. In short, there are two R scripts which work jointly: [**bacen_url**]('https://raw.githubusercontent.com/paulo-icaro/Bacen_API/main/bacen_url.R') e [**bacen_api**]('https://raw.githubusercontent.com/paulo-icaro/Bacen_API/main/bacen_api.R'). The script [**bacen_url**]('https://raw.githubusercontent.com/paulo-icaro/Bacen_API/main/bacen_url.R') is responsible for creating the URL where the data is available. The [**bacen_api**]('https://raw.githubusercontent.com/paulo-icaro/Bacen_API/main/bacen_api.R') connects with the API, extracts the info required and converts it to a readable format. 
|       Additionally, the library provides the [**bacen_query**]('https://raw.githubusercontent.com/paulo-icaro/Bacen_API/main/bacen_query.R') function, which combines the functionality of both previous functions into a single interface, simplifying the data retrieval process.
|       The following sections describe how each function works.
</p>



<!------------>
<!-- PART 2 -->
<!------------>
## Bacen_URL

{{< include includes/bacen_url.qmd >}}

<p>
|       The mechanism is very simple. The user just have to specify three arguments, which are: **series_code**, **start_date**, **end_date**. These arguments correspond, respectively to the series code, the start and the end date. By specifying these arguments the url's is generated.
|       In the example bellow, we are considering the Brazilian price index time series (IPCA), whose code is 433, in the 2010/23 interval. Once you called the function, just write the parameters and the job is done.

**Attention !**

- Have in mind that the date pattern is the brazilian one (*dd/mm/yyyy*);
- Dot not forget to write the dates in string/character format;
</p>


```{r}
#| label: Bacen_URL
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| comment: ''
#| eval: TRUE

# -------------------------------- #
# --- Example - URL Generation --- #
# -------------------------------- #

# --- Bacen_URL Function --- #
source('https://raw.githubusercontent.com/paulo-icaro/Bacen_API/main/bacen_url.R')    

# --- Create URL --- #
ipca_br_url = bacen_url(433, '01/01/2010', '31/12/2023')    # Generating the URL
ipca_br_url
```



<!-------------->
<!--- PART 3 --->
<!-------------->
## Bacen_API

{{< include includes/bacen_api.qmd >}}

<p>
|       This function connects with Bacen API by using the [*httr*](https://httr.r-lib.org/) or [*httr2*](https://httr2.r-lib.org/) functions. In both cases, the [**Bacen_API**](https://raw.githubusercontent.com/paulo-icaro/Bacen_API/main/bacen_api.R) function internally checks the HTTP status code. Put in to other words, it verifies whether we get a successful connection (code 200) or a bad connection (code 400 or 404). In case the initial try fails, there are three remaining chances to achieve a valid connection. Some messages are displayed during the download of the data.
|       The arguments of the function are the ***url*** and a logical variable named ***httr*** indicating whether the user wants do use [*httr*](https://httr.r-lib.org/) or [*httr2*](https://httr2.r-lib.org/). The next example shows how to use the function.
</p>

```{r}
#| label: Bacen_API
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| comment: ''
#| eval: TRUE

# -------------------------------- #
# --- Example - API Connection --- #
# -------------------------------- #

# --- Bacen_API Function --- #
source('https://raw.githubusercontent.com/paulo-icaro/Bacen_API/main/bacen_api.R')

# --- URL --- #
ipca_br_url = 'https://api.bcb.gov.br/dados/serie/bcdata.sgs.433/dados?formato=json&dataInicial=01/01/2003&dataFinal=31/12/2023'

# --- Acessing API Data --- #
data = bacen_api(url = ipca_br_url, httr = TRUE)
tail(data, n = 10)
```
<!-------------->
<!--- PART 4 --->
<!-------------->
## Bacen_Query

{{< include includes/bacen_query.qmd >}}

<p>
|       This script concatenates both former functions. Besides the already known arguments from [**bacen_api**]('https://raw.githubusercontent.com/paulo-icaro/Bacen_API/main/bacen_api.R') function, the researcher only has to inform two more items in order to proceed with the data extraction. First, one needs to inform the ***series_name*** and also decide whether sourcing the **bacen_url** and **bacen_api** functions from this library.
|       Let's practice the data extraction:
</p>

```{r}
#| label: Bacen_Query
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| comment: ''
#| eval: TRUE


# ----------------------- #
# --- Data Extraction --- #
# ----------------------- #

# --- Bacen API Function --- #
source('https://raw.githubusercontent.com/paulo-icaro/Bacen_API/main/bacen_query.R')

# --- Previous Info --- #
cod_bacen_series = c('25390', '4189', '3697', '433')
name_bacen_series = c('ibcrce', 'selic_acum_anual', 'tx_cambio_mp', 'inflação_ipca')
start_date = '01/01/2015'
end_date = '31/12/2025'

# --- Extraction --- #
bacen_dataset = bacen_query(cod_bacen_series, name_bacen_series, start_date, end_date)
tail(bacen_dataset)
```